\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}

%opening
\title{Project Description}
\author{Samuel Pearce}

\begin{document}

\maketitle

\tableofcontents

\pagebreak

\section{Problem Description}
The problem to be inquired into as the subject of this final paper (VA)
is that the acquisition of language is still a poorly-understood process.
The main problem is also that I am unaware of the true extent of this field.
Through this VA, I intend to expand my knowledge on the subject and conduct
a small, informal experiment of my own, to document the process of language
acquisition through immersion in comprehensible input.

\section{Current State of Research}
At the moment, the most commonly accepted --- but still highly controversial
--- theory, is that humans have an inate biological framework for language
that allows us to acquire it through immersion when we are extremely young.
This implies the existence of some structure that must be common to all of
the world's languages --- what's known as Chomsky's universal grammar. This
has prompted the research and documentation of many languages and their base
structures to find the common thread.

So far, that common feature seems to be recursion --- the ability to build
on top of previous ideas to create succesively complex concepts. This is
most obviously seen in languages where multiple sentences can be nested. For
example: ``The man, who's car had been rusting, because it had rained a lot,
liked to eat fish.''

There has been some evidence that might debunk this claim, but more samples
of languages are needed and much more time is required to process it in order to
conclusively prove any theory of universal grammar.

\section{Questions to be Answered}
There are two questions to be answered by myself (Samuel Pearce) by this VA.
They are as follows:
\begin{enumerate}
 \item What, if any, linguistic features are univeral to all natural human languages,
 and what are the known facts about Chomsky's universal grammar, as of the time of this VA?
 \item How does language acquisition work in adults and how effectively can an
 adult learn a language through immersion and comprehensible input alone?
\end{enumerate}
The goal of the final document to be handed in, is to provide answers to these
questions to the best of my ability and knowledge. The final document must also
cite all utilised sources for the claims made therein.

\section{Method}
The method I have chosen to answer these questions is twofold:

For the first part --- the UG Summary --- I will research universal grammar by
reading one (or two, if time permits) linguistic textbooks on the subject and write a summary of their
contents while citing the sources of my claims. The aim of the summary is to show
my understanding of the topic and prove my ability to use academic sources in a
paper. The summary will be written as a surface-level introduction aimed towards
readers who are familiar with the fundamentals of linguistics.

The second part --- the experiment --- will be used to gain information on the
acquisition of language through pure immersion for adults.
I will spend time with volunteers speaking toki pona
\footnote{
toki pona is a minimal language constructed by Sonja Lang in 2001. It is
designed to have as small of a lexicon as possible, while still being able to express
complex concepts. It was created in order to simplify the speaker's thoughts, according
to the Sapir-Whorf hypothesis.
}
This language has been chosen for its simple grammar
and lexicon size, in order to keep the time-frame of the experiment minimal. One of the
volunteers will be provided with lessons and flashcards to learn the language with, the
``traditional'' way, while the other will be simply immersed in the language with me
answering any questions in the language and provided with pure toki pona resources. Once
the volunteers have reached a minimum of semi-fluency (the point where the most basic
conversations can be had), I will record an interview with them in toki pona to show
their ability. The interview will be transcribed and translated into a report for the
final document. This report will also include details of the acquisition process and
my method of teaching.

Once both halves of the project have been completed, the UG Summary, the Experiment report,
and a journal of the steps taken throughout the whole project (the Work Log) will
be compiled with introduction and review sections into the final report.

\section{Required Materials}
For required materials, I have listed some books which should serve as a good
introduction as well as solid literary source for the UG Summary. They are:
\begin{itemize}
 \item Chomsky's Universal Grammar: An Introduction\\(ISBN-13: 978-1-4051-1186-7)
 \item Universal Grammar in Second-Language Acquisition: A History\\(ISBN-13: 978-0-5217-9647-7)
\end{itemize}

For the Experiment, I will use ``Toki Pona: The Language of Good (ISBN-13: 978-0-9782-9230-0)''
as the reference grammar for toki pona.
For the interview with the volunteers, I will use audio recording software such as Audacity,
or the audio recording capabilities of my phone. All documentation will either be performed
using Microsoft Word or PDFLaTeX and Kile editor.

\section{Task Division}
The tasks can, for the most part, be run in parallel; the UG Summary can be researched and
written during the time provided from TBZ, and the experiment can be run in my free time,
so that I can conform to the schedules of the volunteers. This way, both parts of the project
can be compeleted at the same time before compiling the results into the final document.

The task division for the UG summary would only really consist of first reading through
the selected literature to look for the most important points to compile in the sumamry,
which would be followed by writing the summary while citing the sources. Finally, the
written summary would be redrafted and tweaked according to feedback.

For the experiment, the first part would consist of teaching the volunteers toki pona
until time runs out, or they reach a desired level of fluency for an interview. After
which, the interview would be performed and documented.

\section{Time Plan}
I propose to divide the allotted four weeks as follows:
\begin{itemize}
 \item \textbf{Week 1:}
 \begin{itemize}
  \item Read selected UG literature
  \item Provide volunteers with comprehensible input
 \end{itemize}
\item \textbf{Week 2:}
 \begin{itemize}
  \item Write first draft of UG Summary with sources
  \item Provide volunteers with comprehensible input
 \end{itemize}
\item \textbf{Week 3:}
 \begin{itemize}
  \item Second, Third, etc. drafts of UG Summary with suggested changes
  \item Provide volunteers with comprehensible input
 \end{itemize}
\item \textbf{Week 4:}
 \begin{itemize}
  \item Finalise UG Summary
  \item Interview volunteers and transcribe/translate interview
  \item Compile results in final document
 \end{itemize}
\end{itemize}

\section{Signatures}
\hfill \break
\hfill \break
\noindent\begin{tabular}{ll}
\makebox[2.5in]{\hrulefill} & \makebox[2.5in]{\hrulefill}\\
Student & Date\\[8ex]% adds space between the two sets of signatures
\makebox[2.5in]{\hrulefill} & \makebox[2.5in]{\hrulefill}\\
Teacher & Date\\
\end{tabular}

\end{document}
